% ============================================================================
% pygeoinf Technical Manual
% Theory, Algorithms, and API Reference
% ============================================================================
\documentclass[11pt,twoside,openright]{book}

% ---------------------------------------------------------------------------
% Typography and layout
% ---------------------------------------------------------------------------
\usepackage[final]{microtype}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage[margin=1in,headheight=14pt]{geometry}
\usepackage{fancyhdr}
\usepackage{titlesec}
\usepackage{parskip}          % paragraph spacing instead of indent

% ---------------------------------------------------------------------------
% Mathematics
% ---------------------------------------------------------------------------
\usepackage{amsmath,amssymb,amsthm}
\usepackage{mathtools}
% \usepackage{physics}  % install texlive-science if needed
\usepackage{bm}

% ---------------------------------------------------------------------------
% References, links, cross-references
% ---------------------------------------------------------------------------
\usepackage[hidelinks,bookmarksnumbered]{hyperref}
\usepackage[nameinlink,capitalise]{cleveref}
\usepackage[numbers,sort&compress]{natbib}

% ---------------------------------------------------------------------------
% Figures, diagrams, tables
% ---------------------------------------------------------------------------
\usepackage{graphicx}
\usepackage{tikz-cd}
\usepackage{tikz}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{booktabs}
\usepackage{array}
\usepackage{longtable}

% ---------------------------------------------------------------------------
% Code and algorithm environments
% ---------------------------------------------------------------------------
\usepackage{listings}
\usepackage[ruled,vlined]{algorithm2e}

% ---------------------------------------------------------------------------
% Boxed callouts for code links
% ---------------------------------------------------------------------------
\usepackage[framemethod=tikz]{mdframed}

\newmdenv[
  linewidth=0.6pt,
  roundcorner=4pt,
  backgroundcolor=blue!3,
  linecolor=blue!30,
  skipabove=\medskipamount,
  skipbelow=\medskipamount,
  frametitle={\small\sffamily pygeoinf API},
  frametitlebackgroundcolor=blue!10,
  frametitlerule=true,
  frametitlerulewidth=0.4pt,
]{codebox}

\newmdenv[
  linewidth=0.6pt,
  roundcorner=4pt,
  backgroundcolor=yellow!5,
  linecolor=yellow!50!orange,
  skipabove=\medskipamount,
  skipbelow=\medskipamount,
  frametitle={\small\sffamily Why this matters},
  frametitlebackgroundcolor=yellow!15,
  frametitlerule=true,
  frametitlerulewidth=0.4pt,
]{whybox}

\newmdenv[
  linewidth=0.6pt,
  roundcorner=4pt,
  backgroundcolor=green!3,
  linecolor=green!40,
  skipabove=\medskipamount,
  skipbelow=\medskipamount,
  frametitle={\small\sffamily Worked Example},
  frametitlebackgroundcolor=green!10,
  frametitlerule=true,
  frametitlerulewidth=0.4pt,
]{examplebox}

\newmdenv[
  linewidth=0.6pt,
  roundcorner=4pt,
  backgroundcolor=red!3,
  linecolor=red!30,
  skipabove=\medskipamount,
  skipbelow=\medskipamount,
  frametitle={\small\sffamily Common Pitfall},
  frametitlebackgroundcolor=red!8,
  frametitlerule=true,
  frametitlerulewidth=0.4pt,
]{pitfallbox}

% ---------------------------------------------------------------------------
% Theorem environments
% ---------------------------------------------------------------------------
\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[chapter]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}

\theoremstyle{remark}
\newtheorem*{remark}{Remark}

% ---------------------------------------------------------------------------
% Custom commands: Spaces and maps
% ---------------------------------------------------------------------------
\newcommand{\modelspace}{\mathcal{M}}
\newcommand{\dataspace}{\mathcal{D}}
\newcommand{\propertyspace}{\mathcal{P}}
\newcommand{\Gd}{\mathbf{G}}              % standard discrete forward matrix
\newcommand{\GdGal}{\tilde{\mathbf{G}}}   % Galerkin discrete matrix
\newcommand{\Tau}{\mathcal{T}}             % property operator

% ---------------------------------------------------------------------------
% Custom commands: Measures
% ---------------------------------------------------------------------------
\newcommand{\prior}{\mu_{\modelspace}^0}
\newcommand{\post}{\mu_{\modelspace}^{\mathbf{\tilde{d}}}}
\newcommand{\discprior}{\mu_{\modelspace}^{0,N}}
\newcommand{\discpost}{\mu_{\modelspace}^{\mathbf{\tilde{d}}, N}}

% ---------------------------------------------------------------------------
% Custom commands: Discretization
% ---------------------------------------------------------------------------
\newcommand{\Vspace}{\mathbb{V}_N}
\newcommand{\mass}{\mathbf{M}}             % Gram/mass matrix

% ---------------------------------------------------------------------------
% Custom commands: Inner products and pairings
% ---------------------------------------------------------------------------
\newcommand{\ipM}[2]{(#1,#2)_{\modelspace}}
\newcommand{\ipV}[2]{(#1,#2)_{\Vspace}}
\newcommand{\ipD}[2]{(#1,#2)_{\dataspace}}
\newcommand{\ipH}[2]{(#1,#2)_{H}}
\newcommand{\ip}[3]{(#1,#2)_{#3}}         % generic inner product
\newcommand{\pair}[2]{\langle #1, #2\rangle}

% ---------------------------------------------------------------------------
% Custom commands: Riesz maps and adjoint
% ---------------------------------------------------------------------------
\newcommand{\RieszM}{\mathcal{R}_{\modelspace}}
\newcommand{\RieszD}{\mathcal{R}_{\dataspace}}
\newcommand{\Riesz}{\mathcal{R}}
\newcommand{\Gadj}{G^*}

% ---------------------------------------------------------------------------
% Custom commands: Math operators
% ---------------------------------------------------------------------------
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator{\spn}{span}
\DeclareMathOperator{\supp}{supp}
\DeclareMathOperator{\dom}{dom}
\DeclareMathOperator{\epi}{epi}
\DeclareMathOperator{\diag}{diag}
\DeclareMathOperator{\tr}{tr}
\DeclareMathOperator{\rank}{rank}
\DeclareMathOperator{\range}{range}
\DeclareMathOperator{\kernel}{ker}

% ---------------------------------------------------------------------------
% Listings style for Python
% ---------------------------------------------------------------------------
\lstdefinestyle{pygeoinf}{
  language=Python,
  basicstyle=\small\ttfamily,
  keywordstyle=\color{blue!70!black}\bfseries,
  commentstyle=\color{green!50!black}\itshape,
  stringstyle=\color{red!60!black},
  showstringspaces=false,
  breaklines=true,
  frame=none,
  xleftmargin=0.5em,
  aboveskip=0.5em,
  belowskip=0.5em,
}
\lstset{style=pygeoinf}

% ---------------------------------------------------------------------------
% Page headers and footers
% ---------------------------------------------------------------------------
\pagestyle{fancy}
\fancyhf{}
\fancyhead[LE]{\small\slshape\nouppercase{\leftmark}}
\fancyhead[RO]{\small\slshape\nouppercase{\rightmark}}
\fancyfoot[C]{\thepage}
\renewcommand{\headrulewidth}{0.4pt}

% ---------------------------------------------------------------------------
% Title information
% ---------------------------------------------------------------------------
\title{%
  \vspace{-2cm}
  {\Huge\bfseries pygeoinf}\\[0.8em]
  {\LARGE Technical Manual}\\[0.4em]
  {\Large Theory, Algorithms, and API Reference}\\[2em]
  \rule{\textwidth}{0.8pt}
}
\author{%
  Adrian\\[0.5em]
  {\normalsize Department of Earth Sciences}
}
\date{\today}

% ============================================================================
\begin{document}
% ============================================================================

\frontmatter

% ---------------------------------------------------------------------------
% Title page
% ---------------------------------------------------------------------------
\maketitle

% ---------------------------------------------------------------------------
% Preface
% ---------------------------------------------------------------------------
\chapter{Preface}

\section*{Who this manual is for}

This manual is written for geophysicists who have a forward operator and
a dataset and want to perform rigorous linear inversion or inference. The
word \emph{rigorous} carries weight here: pygeoinf does not cut corners.
It insists on correct Hilbert space structure, proper adjoints, and
treatment of uncertainty that is mathematically honest. If you are used
to inverting $( \mathbf{G}^\top \mathbf{G})^{-1}\mathbf{G}^\top$
and calling it done, this manual will explain why that is sometimes
wrong, and what to do instead.

No prior knowledge of functional analysis is assumed, but you will need
to engage with the mathematics. The package deliberately surfaces
conceptual machinery --- Riesz maps, dual pairings, Hilbert adjoints ---
that many inversion codes hide. The reward is a library whose results are
correct by construction for \emph{any} choice of model space, inner
product, and forward operator, not just the Euclidean special case.

\section*{How to read this manual}

Each chapter follows a consistent four-part structure:
\begin{enumerate}
  \item \emph{Why this matters} --- a motivating argument, usually with a
    geophysical example, that explains why the concept is worth the
    effort. Highlighted in yellow callout boxes.
  \item \emph{Definitions and theorems} --- the full, rigorous mathematical
    treatment.
  \item \emph{Implementation in pygeoinf} --- how each mathematical concept
    maps to the API. Highlighted in blue callout boxes labelled
    ``pygeoinf API''.
  \item \emph{Worked example} --- a complete mini-example, in both
    mathematics and code. Highlighted in green callout boxes.
\end{enumerate}

If you are primarily a practitioner, you can read the yellow and blue
boxes and skip the proofs on a first pass without losing the ability to
use the package.
If you are mathematically inclined, read everything; proofs that would
interrupt the flow are deferred to \cref{app:proofs}.

Red callout boxes labelled ``Common Pitfall'' flag mistakes that are
easy to make and important to avoid.

\section*{Acknowledgements}

This manual documents work carried out at the Department of Earth
Sciences. The theoretical framework draws heavily on
\citet{al2021linear}, \citet{stuart2010inverse}, and the trilogy
\citet{backus1970inference,backus1970uniqueness,backus1970inferenceIII}.

% ---------------------------------------------------------------------------
% Table of contents
% ---------------------------------------------------------------------------
\tableofcontents

% ---------------------------------------------------------------------------
% List of notation
% ---------------------------------------------------------------------------
\chapter{Notation and Conventions}
\label{notation}

\section*{General conventions}

Throughout this manual we use the following typographic conventions.

\begin{itemize}
  \item \textbf{Boldface roman} ($\mathbf{m}$, $\mathbf{G}$, $\boldsymbol{\lambda}$)
    denotes finite-dimensional vectors and matrices, i.e.\ elements of
    $\mathbb{R}^n$ or linear maps $\mathbb{R}^n \to \mathbb{R}^m$.
  \item \textbf{Calligraphic} ($\mathcal{M}$, $\mathcal{D}$, $\mathcal{P}$)
    denotes abstract (possibly infinite-dimensional) Hilbert or Banach
    spaces.
  \item \textbf{Script/Roman capitals} ($G$, $T$, $\Pi$, $\pi$, $\mathcal{R}$)
    denote abstract linear operators between such spaces.
  \item Indices: $j,k,\ell$ typically index a basis of the model space;
    $i$ typically indexes a datum.
  \item We work exclusively over $\mathbb{R}$; all spaces are real.
\end{itemize}

\section*{Spaces}

\begin{longtable}{lll}
\toprule
Symbol & Meaning & Code \\
\midrule
$\modelspace$ & Model space (Hilbert) & \texttt{model\_space} \\
$\dataspace$ & Data space (usually $\mathbb{R}^{N_d}$) & \texttt{data\_space} \\
$\mathcal{P}$ & Property space (usually $\mathbb{R}^{N_p}$) & \texttt{property\_space} \\
$\modelspace'$ & Continuous dual of $\modelspace$ & implicit via Riesz \\
$\dataspace'$ & Continuous dual of $\dataspace$ & implicit via Riesz \\
$H$ & Generic Hilbert space & \texttt{HilbertSpace} \\
$\ell^2$ & Square-summable sequences & infinite-dim.~limit \\
$\mathbb{R}^N$ & Finite coefficient space & \texttt{EuclideanSpace(N)} \\
$\modelspace_N$ & $N$-dimensional subspace of $\modelspace$ & \texttt{dim = N} \\
$H_1 \oplus H_2$ & Hilbert direct sum & \texttt{HilbertSpaceDirectSum} \\
\bottomrule
\end{longtable}

\section*{Operators and maps}

\begin{longtable}{lll}
\toprule
Symbol & Meaning & Code \\
\midrule
$G : \modelspace \to \dataspace$ & Forward operator & \texttt{forward\_operator} \\
$G' : \dataspace' \to \modelspace'$ & Banach dual (transpose) & \texttt{.dual} \\
$G^* : \dataspace \to \modelspace$ & Hilbert adjoint & \texttt{.adjoint} \\
$\mathcal{T} : \modelspace \to \mathcal{P}$ & Property operator & \texttt{property\_operator} \\
$\mathcal{R}_X : X' \to X$ & Riesz isomorphism & \texttt{from\_dual} \\
$\mathcal{R}_X^{-1} : X \to X'$ & Inverse Riesz & \texttt{to\_dual} \\
$\Pi : \modelspace \to \ell^2$ & Analysis (expansion coefficients) & \texttt{to\_components} \\
$\pi : \ell^2 \to \modelspace$ & Synthesis (reconstruct from coeffs) & \texttt{from\_components} \\
$\Pi_N : \modelspace \to \mathbb{R}^N$ & Truncated analysis & \texttt{to\_components} (finite) \\
$\pi_N : \mathbb{R}^N \to \modelspace_N$ & Truncated synthesis & \texttt{from\_components} (finite) \\
$P_N : \modelspace \to \modelspace_N$ & Orthogonal projection onto $\modelspace_N$ & implicit \\
$\phi_j \in \modelspace$ & Basis vector & \texttt{basis\_vector(j)} \\
$\psi_j \in \modelspace$ & Biorthogonal dual basis vector & derived from Gram \\
\bottomrule
\end{longtable}

\section*{Inner products and pairings}

\begin{longtable}{lll}
\toprule
Symbol & Meaning & Code \\
\midrule
$(m_1, m_2)_{\modelspace}$ & Inner product on $\modelspace$ & \texttt{inner\_product(x1,x2)} \\
$(d_1, d_2)_{\dataspace}$ & Inner product on $\dataspace$ & \texttt{inner\_product(d1,d2)} \\
$\langle \ell, m \rangle$ & Dual pairing: $\ell \in \modelspace',\, m \in \modelspace$ & \texttt{l(m)} \\
$\|m\|_{\modelspace}$ & Hilbert norm $\sqrt{(m,m)_{\modelspace}}$ & \texttt{norm(m)} \\
\bottomrule
\end{longtable}

We use \emph{parentheses} $(\cdot,\cdot)$ for inner products and
\emph{angle brackets} $\langle\cdot,\cdot\rangle$ for dual pairings.
On a Hilbert space the two are related by
$\langle \ell, m \rangle = (m,\, \mathcal{R}_X \ell)_X$.

\section*{Matrices (finite-dimensional objects)}

\begin{longtable}{lll}
\toprule
Symbol & Meaning & Notes \\
\midrule
$\mathbf{M}_{\phi,N}$ & Gram/mass matrix $[\,(\phi_k,\phi_j)_{\modelspace}\,]_{j,k}$ & SPD, $N\times N$ \\
$\mathbf{G}$ & Standard discrete matrix of $G$: $[\,(G\phi_j,\beta_i)_{\dataspace}\,]$ & $N_d \times N$ \\
$\tilde{\mathbf{G}}$ & Galerkin matrix (with inner products) & \texttt{.matrix(galerkin=True)} \\
$\mathbf{m}, \mathbf{d}$ & Coefficient vectors in $\mathbb{R}^N$, $\mathbb{R}^{N_d}$ & \texttt{to\_components(m)} \\
\bottomrule
\end{longtable}

\section*{Measures and probability}

\begin{longtable}{lll}
\toprule
Symbol & Meaning & Code \\
\midrule
$\mu$ & A probability measure on $\modelspace$ & \\
$\mathcal{N}(\mu_0, C)$ & Gaussian measure, mean $\mu_0$, covariance $C$ & \texttt{GaussianMeasure} \\
$\mu^0$ & Prior measure & \texttt{prior} \\
$\mu^{\tilde{d}}$ & Posterior measure given data $\tilde{d}$ & \texttt{.invert(data)} \\
$C : \modelspace \to \modelspace$ & Covariance operator & \texttt{.covariance\_operator} \\
$\mathbf{C}_{\dataspace} : \dataspace \to \dataspace$ & Data noise covariance & \texttt{data\_error\_measure} \\
$\varepsilon \in \dataspace$ & Data noise, $\varepsilon \sim \mathcal{N}(0,\mathbf{C}_{\dataspace})$ & \\
\bottomrule
\end{longtable}

\section*{Convex analysis}

\begin{longtable}{lll}
\toprule
Symbol & Meaning & Code \\
\midrule
$C \subset \modelspace$ & A convex subset & \texttt{Subset} subclass \\
$h_C(q) = \sigma_C(q)$ & Support function of $C$ & \texttt{SupportFunction} \\
$\partial h_C(q)$ & Subdifferential of $h_C$ at $q$ & \texttt{.subgradient(q)} \\
$\mathcal{U} \subset \mathcal{P}$ & Admissible property set & computed by DLI \\
\bottomrule
\end{longtable}

% ============================================================================
\mainmatter
% ============================================================================

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\part{Foundations}
\label{part:foundations}
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% ============================================================================
\chapter{Introduction and Motivation}
\label{ch:introduction}
% ============================================================================

\section{What is pygeoinf?}
\label{sec:what-is-pygeoinf}

pygeoinf is a Python library for rigorous linear geophysical inversion and
Bayesian inference. It is built around a single design philosophy:
\emph{the mathematical structure of the problem should be encoded explicitly
in the code}, not reduced to bare matrices and vectors.

Most inversion codes operate on the discrete level from the start: the
user assembles a matrix $\mathbf{G}\in\mathbb{R}^{M\times N}$, a data
vector $\mathbf{d}\in\mathbb{R}^M$, and calls some solver. This approach
works, but it forces the user to manually track which inner products
are implicit, which transposes are adjoints, and how uncertainty
propagates through the discretization. When anything non-standard
happens --- a non-Euclidean norm, a non-uniform grid, a Sobolev
prior --- the bookkeeping becomes error-prone.

Instead, pygeoinf exposes the following conceptual objects as first-class
Python types:

\begin{itemize}
  \item \textbf{Hilbert spaces} (\texttt{HilbertSpace}) --- carrying their
    inner product, Riesz maps, and basis structure.
  \item \textbf{Linear operators} (\texttt{LinearOperator}) --- aware of
    their domain and codomain, automatically computing the correct Hilbert
    adjoint from the spaces' Riesz maps.
  \item \textbf{Linear forms} (\texttt{LinearForm}) --- elements of the dual
    space, representing individual measurements.
  \item \textbf{Gaussian measures} (\texttt{GaussianMeasure}) --- priors and
    posteriors as probability measures on Hilbert spaces.
  \item \textbf{Convex sets and support functions} --- geometry of the
    admissible model set.
\end{itemize}

Once these objects are defined, a full suite of inversion algorithms ---
Bayesian, least-squares, minimum-norm, Backus-Gilbert, and
deterministic linear inference --- work automatically and correctly
for any choice of spaces and operators.

\section{The geophysical inversion problem}
\label{sec:geophysical-inversion}

The starting point is a physical system described by a \emph{model}
$m$ --- perhaps the seismic wave-speed field $v_P(x)$, the electrical
conductivity $\sigma(x)$, or the subsurface density $\rho(x)$. Some
aspects of this system are measurable: we collect \emph{data}
$\tilde{d} = (\tilde{d}_1,\ldots,\tilde{d}_M)$ via instruments
(travel-time picks, gravity readings, magnetic measurements, \ldots).
A \emph{forward model} $G$ maps a model to its predicted data:
\begin{equation}
  \label{eq:forward-model}
  d = Gm + \varepsilon,
\end{equation}
where $\varepsilon$ is measurement noise. The \emph{inverse problem} is
to determine what equation~\eqref{eq:forward-model} implies about
$m$ given the observed data $\tilde{d}$.

All realistic geophysical inverse problems share three features:

\begin{enumerate}
  \item \textbf{Underdetermination.} The data $\tilde{d}$ is a
    finite-dimensional vector, whereas $m$ is a function (infinite-dimensional
    in principle). Infinitely many models are consistent with the data
    exactly; the problem has a non-trivial null space.
  \item \textbf{Ill-conditioning.} Even the components of $m$ that \emph{are}
    constrained by the data may be constrained very weakly, amplifying
    noise in the solution.
  \item \textbf{Incomplete prior knowledge.} We have partial information
    about $m$ --- physical bounds, smoothness expectations, perhaps a
    statistical prior --- that should inform the inversion.
\end{enumerate}

Any honest inversion framework must confront all three. pygeoinf does so
by working at the level of \emph{abstract spaces and operators}, which
makes the structure of the problem explicit and prevents silent,
basis-dependent errors.

\section{Why abstract spaces matter}
\label{sec:why-spaces-matter}

\begin{whybox}
If you are used to working with matrices and vectors in NumPy, you may
wonder why pygeoinf insists on Hilbert spaces, dual pairings, and Riesz
maps. This section shows you exactly what goes wrong when you ignore
these structures --- and how pygeoinf gets it right automatically.
\end{whybox}

To make the argument concrete, consider a simple 1D problem. You want to
recover a function $m : [0,1] \to \mathbb{R}$ from $M$ integral
measurements
\begin{equation}
  \label{eq:motivating-forward}
  d_i = \int_0^1 g_i(x)\, m(x)\, \mathrm{d}x, \qquad i = 1,\ldots,M.
\end{equation}
You discretize $[0,1]$ into $N$ cells with centres $x_j^*$ and widths
$\Delta x_j > 0$ (possibly non-uniform). Approximate $m$ as piecewise
constant: $m(x) \approx m_j$ on cell $j$, so $\mathbf{m} =
(m_1,\ldots,m_N)^\top \in \mathbb{R}^N$. The discrete forward matrix
becomes
\begin{equation}
  \label{eq:motivating-matrix}
  G_{ij} = g_i(x_j^*) \Delta x_j,
  \qquad \mathbf{d} = \mathbf{G}\mathbf{m} \in \mathbb{R}^M.
\end{equation}
You want the minimum-norm solution: among all $\mathbf{m}$ with
$\mathbf{G}\mathbf{m} = \mathbf{d}$, find the one of smallest
norm.

\paragraph{The naive approach: use the Euclidean norm on $\mathbb{R}^N$.}
Minimizing $\|\mathbf{m}\|_2^2 = \sum_j m_j^2$ gives
\begin{equation}
  \label{eq:naive-min-norm}
  \mathbf{m}^\dagger_{\text{naive}}
  = \mathbf{G}^\top (\mathbf{G}\mathbf{G}^\top)^{-1} \mathbf{d}.
\end{equation}

\paragraph{The correct approach: use the $L^2([0,1])$ norm.}
The natural inner product approximating $\|m\|_{L^2}^2 = \int_0^1 m(x)^2
\,\mathrm{d}x$ is
\begin{equation}
  \label{eq:L2-inner-product}
  (\mathbf{m},\mathbf{n})_{L^2} \approx \sum_j m_j n_j \Delta x_j
  = \mathbf{m}^\top \mathbf{M} \mathbf{n},
  \qquad \mathbf{M} = \mathrm{diag}(\Delta x_1,\ldots,\Delta x_N).
\end{equation}
Minimizing this gives the qualitatively different solution
\begin{equation}
  \label{eq:correct-min-norm}
  \mathbf{m}^\dagger
  = \mathbf{M}^{-1}\mathbf{G}^\top
    (\mathbf{G}\mathbf{M}^{-1}\mathbf{G}^\top)^{-1} \mathbf{d}.
\end{equation}

\paragraph{Why they differ.}
If the grid is uniform ($\Delta x_j = 1/N$ for all $j$), then
$\mathbf{M} = \tfrac{1}{N}\mathbf{I}$ and the two solutions agree up to
a constant. But on a non-uniform grid, the naive solution
\eqref{eq:naive-min-norm} \emph{implicitly weights finely-sampled
regions more heavily}: minimizing $\sum_j m_j^2$ without the $\Delta x_j$
weights is not minimizing a physically meaningful norm of $m$.

The matrix $\mathbf{M}$ is the \emph{Gram matrix} (or mass matrix) of the
basis $\{\phi_j\}$ where $\phi_j = \mathbf{1}_{\text{cell }j} / \Delta x_j^{1/2}$.
The formula $G^* = \mathbf{M}^{-1}\mathbf{G}^\top$ is the finite-dimensional
instance of the abstract relation
\begin{equation}
  \label{eq:hilbert-adjoint-abstract}
  G^* = \mathcal{R}_{\modelspace} \circ G' \circ \mathcal{R}_{\dataspace}^{-1},
\end{equation}
which pygeoinf enforces automatically once you provide the spaces with
their correct inner products.

\begin{pitfallbox}
The matrix transpose $\mathbf{G}^\top$ is the adjoint only in Euclidean
space with the standard dot product. In any other Hilbert space it is
not. Using $\mathbf{G}^\top$ instead of
$\mathbf{M}_{\mathcal{M}}^{-1}\mathbf{G}^\top\mathbf{M}_{\mathcal{D}}$
produces a solution that is biased toward densely-gridded regions (or,
in a Bayesian setting, posterior covariances that are not
basis-independent).
\end{pitfallbox}

This example is the simplest case. As problems become more complex ---
Sobolev regularity norms, non-trivial data weighting, operators between
function spaces --- the gap between ``naive'' and ``correct'' widens.
pygeoinf closes that gap structurally, not by asking the user to remember
the right formula each time.

\section{Package architecture overview}
\label{sec:architecture}

pygeoinf is organized in four ascending layers.

\medskip
\begin{center}
\begin{tikzpicture}[
  layer/.style={draw, rounded corners=4pt, fill=#1, minimum width=8cm,
                minimum height=0.75cm, font=\small\sffamily},
  arr/.style={->, thick, gray!60}
]
  \node[layer=blue!10]   (L1) at (0,0)    {Foundation: spaces, operators, linear forms};
  \node[layer=green!10]  (L2) at (0,1.0)  {Geometry: subsets, subspaces, support functions, measures};
  \node[layer=yellow!15] (L3) at (0,2.0)  {Algorithms: forward problems, inversions, optimisation};
  \node[layer=red!8]     (L4) at (0,3.0)  {Output: visualisation, diagnostics};
  \draw[arr] (L1) -- (L2);
  \draw[arr] (L2) -- (L3);
  \draw[arr] (L3) -- (L4);
\end{tikzpicture}
\end{center}
\medskip

\paragraph{Foundation layer}
(\cref{part:foundations}) The abstract base classes
\texttt{HilbertSpace}, \texttt{LinearOperator}, and \texttt{LinearForm}
define the mathematical objects.  Everything else is built on top of
them.  No inversion algorithm is aware of the concrete space type; it
only calls the abstract interface.

\paragraph{Geometry layer}
(\cref{part:geometry}) Convex subsets (\texttt{Ball},
\texttt{PolyhedralSet}, \ldots), subspaces (\texttt{AffineSubspace}),
support functions (\texttt{SupportFunction}), and Gaussian measures
(\texttt{GaussianMeasure}) all live here.  They each accept any
\texttt{HilbertSpace} and work correctly regardless of its concrete type.

\paragraph{Algorithm layer}
(\cref{part:algorithms}) \texttt{ForwardProblem},
\texttt{LinearBayesianInversion}, \texttt{LinearMinimumNormInversion},
\texttt{BackusGilbertInversion}, and the convex optimisation algorithms
belong here.  They compose the foundation and geometry objects into
complete solution procedures.

\paragraph{Output layer}
(\cref{part:computation}) \texttt{SubspaceSlicePlotter} and related
tools for visualising results and checking algorithmic correctness.

\section{How to read this manual}
\label{sec:how-to-read}

The manual is divided into four parts mirroring the package architecture:

\begin{description}
  \item[\cref{part:foundations} (Foundations)]
    Chapters~\ref{ch:hilbert-spaces}--\ref{ch:nonlinear}.
    The mathematical vocabulary: spaces, operators, forms.
    \emph{All users should read at least the motivation sections of these
    chapters.}
  \item[\cref{part:geometry} (Geometry)]
    Chapters~\ref{ch:subsets}--\ref{ch:gaussian-measures}.
    Convex geometry, subspaces, and probabilistic models.
  \item[\cref{part:algorithms} (Problems and Algorithms)]
    Chapters~\ref{ch:forward-problems}--\ref{ch:convex-optimisation}.
    How to set up and solve inverse problems.
  \item[\cref{part:computation} (Computational Methods)]
    Chapters~\ref{ch:computational}--\ref{ch:visualisation}.
    Numerical backends and visualisation.
\end{description}

Each chapter is self-contained. Forward references are kept to a minimum.
The appendices collect a convex analysis primer
(\cref{app:convex-primer}), functional analysis prerequisites
(\cref{app:functional-analysis}), deferred proofs
(\cref{app:proofs}), and a complete API quick-reference
(\cref{app:api-reference}).

\paragraph{Callout boxes.}
\begin{description}
  \item[\textcolor{yellow!60!gray}{Yellow} --- Why this matters]
    Motivational content and geophysical perspective.
  \item[\textcolor{blue!60}{Blue} --- pygeoinf API]
    The mapping from mathematical concept to Python code.
  \item[\textcolor{green!60!gray}{Green} --- Worked Example]
    Complete mini-examples with both equations and code.
  \item[\textcolor{red!60}{Red} --- Common Pitfall]
    Frequent mistakes and how to avoid them.
\end{description}

\section{Notation and conventions summary}
\label{sec:notation-summary}

A complete notation table is given in the \hyperref[notation]{Notation
and Conventions} chapter of the front matter. The most important rules:

\begin{itemize}
  \item Parentheses $(m_1, m_2)_{\modelspace}$ always denote inner products;
    angle brackets $\langle \ell, m \rangle$ always denote dual pairings.
  \item Boldface ($\mathbf{m}$, $\mathbf{G}$) is reserved for
    finite-dimensional vectors and matrices.
  \item Calligraphic letters ($\modelspace$, $\dataspace$) denote abstract
    spaces; Roman capitals ($G$, $T$, $\Pi$) denote abstract operators.
  \item $G^*$ always means the \emph{Hilbert} adjoint
    ($G^*: \dataspace \to \modelspace$); $G'$ means the \emph{Banach} dual
    ($G': \dataspace' \to \modelspace'$). They are related by
    \eqref{eq:hilbert-adjoint-abstract} but are not the same map.
\end{itemize}


% ============================================================================
\chapter{Hilbert Spaces}
\label{ch:hilbert-spaces}
% ============================================================================

\section{Why Hilbert spaces?}
\label{sec:why-hilbert}

\begin{whybox}
The inner product on your model space is not a minor technical detail ---
it determines what ``similar models'' means, and therefore what your
inversion algorithm considers a good solution. Two inversions that differ
only in their choice of inner product will in general produce qualitatively
different results. pygeoinf makes the inner product an explicit,
first-class object so that this choice is always visible and always
consistent.
\end{whybox}

Why not work with plain vectors and matrices? A vector space gives you
addition and scalar multiplication, but it has no notion of length or
angle. You cannot define a ``smallest'' element, orthogonality, or
projection without more structure. The extra structure needed is an
\emph{inner product}.

For geophysical inversion, the inner product on the model space $\modelspace$
is the mathematical encoding of your regularity expectations about the
model. A Euclidean inner product treats all degrees of freedom equally,
regardless of grid spacing or physical units. An $L^2([0,1])$ inner
product weights by area element $dx$. A Sobolev $H^1$ inner product
additionally penalises gradients. Each choice leads to a different
minimum-norm solution; only the ``right'' one is physically meaningful.

In short: \emph{the inner product is part of the problem specification},
not an implementation detail.

\section{Definition and axioms}
\label{sec:hilbert-definition}

\begin{definition}[Inner product space]
\label{def:inner-product-space}
A \emph{real inner product space} is a real vector space $H$ together with
a map $(\cdot,\cdot)_H : H \times H \to \mathbb{R}$ satisfying, for all
$u, v, w \in H$ and $\alpha \in \mathbb{R}$:
\begin{enumerate}
  \item \textbf{Symmetry:} $(u, v)_H = (v, u)_H$.
  \item \textbf{Linearity in the first argument:}
    $(\alpha u + v, w)_H = \alpha(u, w)_H + (v, w)_H$.
  \item \textbf{Positive definiteness:} $(u, u)_H \geq 0$, with equality only
    if $u = 0$.
\end{enumerate}
\end{definition}

The inner product induces a norm $\|u\|_H := \sqrt{(u, u)_H}$ and hence a
metric $d(u, v) = \|u - v\|_H$.

\begin{definition}[Hilbert space]
\label{def:hilbert-space}
A real \emph{Hilbert space} is a real inner product space $(H, (\cdot,\cdot)_H)$
that is \emph{complete} with respect to the induced norm: every Cauchy
sequence converges to an element of $H$.
\end{definition}

\begin{remark}
pygeoinf works exclusively over $\mathbb{R}$. Complex Hilbert spaces are not
supported. All spaces are assumed to be \emph{separable}, meaning they
admit a countable dense subset; this ensures the existence of countable
orthonormal bases.
\end{remark}

\paragraph{Orthogonality.}
Two elements $u, v \in H$ are \emph{orthogonal}, written $u \perp v$, if
$(u, v)_H = 0$. A set $S \subset H$ is orthogonal if any two distinct
elements of $S$ are orthogonal. If additionally $\|s\|_H = 1$ for all
$s \in S$, the set is \emph{orthonormal}.

\paragraph{The parallelogram law.}
A norm comes from an inner product if and only if it satisfies the
\emph{parallelogram law}:
\begin{equation}
  \label{eq:parallelogram}
  \|u + v\|^2_H + \|u - v\|^2_H = 2\|u\|^2_H + 2\|v\|^2_H
  \qquad \forall\, u, v \in H.
\end{equation}
The $\ell^p$ and $L^p$ spaces for $p \neq 2$ are Banach but not Hilbert;
they do not satisfy \eqref{eq:parallelogram}.

\section{Inner products, norms, and distances}
\label{sec:inner-products-norms}

\begin{proposition}[Cauchy--Schwarz inequality]
\label{prop:cauchy-schwarz}
For any $u, v \in H$,
\begin{equation}
  \label{eq:cauchy-schwarz}
  |(u, v)_H| \leq \|u\|_H \, \|v\|_H,
\end{equation}
with equality if and only if $u$ and $v$ are linearly dependent.
\end{proposition}

The Cauchy--Schwarz inequality guarantees that the cosine of the
``angle'' between two vectors, $\cos\theta_{u,v} := (u,v)_H / (\|u\|_H
\|v\|_H)$, is well-defined and in $[-1,1]$, even in infinite dimensions.

\paragraph{Polarisation identity.}
The inner product can be recovered from the norm via
\begin{equation}
  \label{eq:polarisation}
  (u, v)_H = \tfrac{1}{4}\left(\|u + v\|^2_H - \|u - v\|^2_H\right).
\end{equation}
This means all metric information about a Hilbert space is encoded in
its inner product.

\begin{codebox}
\texttt{HilbertSpace} exposes:
\begin{itemize}
  \item \texttt{inner\_product(x1, x2)}: computes $(x_1, x_2)_H$;
  \item \texttt{norm(x)}: computes $\|x\|_H$;
  \item \texttt{squared\_norm(x)}: computes $\|x\|^2_H$ (often cheaper);
  \item \texttt{distance(x1, x2)}: computes $\|x_1 - x_2\|_H$.
\end{itemize}
Internally, \texttt{inner\_product} is computed via the dual pairing
(see \cref{sec:dual-space}) as $(x_1, x_2)_H = \langle \mathcal{R}^{-1} x_1,\, x_2\rangle$,
using \texttt{to\_dual} and the duality product.
\end{codebox}

\section{The dual space and dual pairing}
\label{sec:dual-space}

\begin{whybox}
In geophysical terms, a dual vector is exactly what an instrument
measures: a number that depends linearly on the model. Understanding the
distinction between vectors (models) and dual vectors (measurements) is
fundamental to understanding why inversion requires Riesz maps.
\end{whybox}

\begin{definition}[Continuous dual space]
\label{def:dual-space}
Let $H$ be a Hilbert space. A \emph{continuous linear functional} on $H$
is a map $\ell : H \to \mathbb{R}$ that is linear ($\ell(\alpha u + v) =
\alpha \ell(u) + \ell(v)$) and bounded ($|\ell(u)| \leq C\|u\|_H$ for
some $C < \infty$). The collection of all continuous linear functionals
on $H$ forms the \emph{(continuous) dual space} $H'$, which is itself a
Banach space under the operator norm $\|\ell\|_{H'} := \sup_{\|u\|=1}
|\ell(u)|$.
\end{definition}

\paragraph{Dual pairing.}
We denote the evaluation of $\ell \in H'$ at $u \in H$ by the \emph{dual
pairing}:
\begin{equation}
  \label{eq:dual-pairing}
  \langle \ell, u \rangle := \ell(u).
\end{equation}
We use angle brackets $\langle\cdot,\cdot\rangle$ for dual pairings and
parentheses $(\cdot,\cdot)_H$ for inner products throughout this manual;
the two are related by the Riesz theorem (\cref{sec:riesz-theorem}).

\paragraph{Two distinct spaces.}
Elements of $H$ are ``primal'' vectors (models); elements of $H'$ are
``dual'' vectors (measurements, functionals, gradients of scalar
functions). On a finite-dimensional Euclidean space, $H \cong H'$ via
the standard dot product, which is why the distinction is often ignored
in numerical practice. On a general Hilbert space --- or more acutely, on
a Banach space where no inner product exists --- the two are different
objects and must be handled separately.

\paragraph{Concrete example.}
Let $H = L^2([0,1])$ and let $g \in L^2([0,1])$. Then the map $\ell:
m \mapsto \int_0^1 g(x) m(x)\,\mathrm{d}x$ is a continuous linear
functional on $H$, i.e.\ $\ell \in H'$. This is precisely the structure
of a geophysical measurement: every seismogram, gravity reading, or
magnetic datum is a dual vector.

\section{The Riesz representation theorem}
\label{sec:riesz-theorem}

\begin{theorem}[Riesz representation]
\label{thm:riesz}
Let $H$ be a real Hilbert space. For every continuous linear functional
$\ell \in H'$ there exists a unique $u_\ell \in H$ such that
\begin{equation}
  \label{eq:riesz-rep}
  \ell(v) = (u_\ell,\, v)_H \qquad \forall\, v \in H.
\end{equation}
Moreover, $\|u_\ell\|_H = \|\ell\|_{H'}$. The map $\Riesz: H' \to H$
defined by $\Riesz(\ell) := u_\ell$ is an isometric isomorphism.
\end{theorem}

\paragraph{Proof sketch.}
If $\ell = 0$ take $u_\ell = 0$. Otherwise, $\ker\ell$ is a closed
proper subspace of $H$ of codimension one. By the projection theorem
there exists a unit vector $e \perp \ker\ell$. Set $u_\ell :=
\ell(e)\,e$. For any $v \in H$, the element $v - \langle\ell,v\rangle e /
\ell(e)$ lies in $\ker\ell$, hence is orthogonal to $e$. A brief
calculation shows $(u_\ell, v)_H = \ell(v)$. Uniqueness follows from
positive definiteness.

\paragraph{Geometric interpretation.}
The Riesz representative $u_\ell = \Riesz(\ell)$ is the element of $H$
that ``points in the direction'' the functional $\ell$ is measuring. In
optimisation this is the \emph{gradient}: if $f(u) = \ell(u)$ then
$\nabla f = u_\ell$.

\paragraph{Riesz map and its inverse.}
We call $\Riesz : H' \to H$ the \emph{Riesz isomorphism} or Riesz map.
Its inverse $\Riesz^{-1} : H \to H'$ is defined by $(\Riesz^{-1} u)(v) =
(u, v)_H$. Together they establish the identification
\begin{equation}
  \label{eq:riesz-iden}
  \langle \ell, v \rangle = (v,\; \Riesz\,\ell)_H
  \qquad \forall\; \ell \in H',\; v \in H.
\end{equation}

\begin{codebox}
The Riesz isomorphism and its inverse are implemented as:
\[
  \texttt{from\_dual} \equiv \Riesz : H' \to H, \qquad
  \texttt{to\_dual}   \equiv \Riesz^{-1} : H \to H'.
\]
If you need them as \texttt{LinearOperator} instances, use the properties
\texttt{HilbertSpace.riesz} ($\Riesz$) and
\texttt{HilbertSpace.inverse\_riesz} ($\Riesz^{-1}$). The dual pairing
evaluates as $\langle \ell, v\rangle = \ell(v)$, implemented by calling
the \texttt{LinearForm} object directly: \texttt{l(v)}.
\end{codebox}

\section{Bases and representations}
\label{sec:bases-representations}

\subsection{Riesz bases and biorthogonal duals}
\label{ssec:riesz-bases}

An \emph{orthonormal basis} (ONB) $\{e_j\}_{j=1}^{\infty}$ is a maximal
orthonormal set; every $m \in H$ expands as $m = \sum_j (m, e_j)_H e_j$.
But ONBs are rarely computable. A physical or numerical basis
$\{\phi_j\}$ is typically not orthonormal, and we must work with
general Riesz bases.

\begin{definition}[Riesz basis]
\label{def:riesz-basis}
A sequence $\{\phi_j\}_{j=1}^{\infty} \subset H$ is a \emph{Riesz basis}
if it is the image of an ONB under a bounded invertible operator. Equivalently,
there exist constants $0 < A \leq B < \infty$ such that for every finite
coefficient sequence $\mathbf{c} \in \mathbb{R}^N$,
\begin{equation}
  \label{eq:riesz-bounds}
  A\, \|\mathbf{c}\|_{\ell^2}^2
  \;\leq\; \Bigl\|\sum_j c_j \phi_j\Bigr\|_H^2
  \;\leq\; B\, \|\mathbf{c}\|_{\ell^2}^2.
\end{equation}
\end{definition}

Every Riesz basis $\{\phi_j\}$ has a unique \emph{biorthogonal dual}
$\{\psi_j\} \subset H$ satisfying
\begin{equation}
  \label{eq:biorthogonality}
  (\phi_j, \psi_k)_H = \delta_{jk}.
\end{equation}
Any element $m \in H$ then has the unique expansion
\begin{equation}
  \label{eq:riesz-expansion}
  m = \sum_{j=1}^{\infty} [\mathbf{m}]_j\, \phi_j,
  \qquad [\mathbf{m}]_j = (m, \psi_j)_H.
\end{equation}
The coefficient sequence $\mathbf{m} = ([\mathbf{m}]_j)_{j=1}^\infty$
lies in $\ell^2$, and the norms are equivalent (eq.~\eqref{eq:riesz-bounds}).
When $\{\phi_j\}$ is an ONB, $\psi_j = \phi_j$ and $A = B = 1$.

\paragraph{Connection to the dual basis.}
The dual basis $\{\phi^i\} \subset H'$ is defined by $\langle \phi^i, \phi_j\rangle
= \delta^i_j$. The biorthogonal dual vectors are the Riesz representatives
of the dual basis:
\begin{equation}
  \label{eq:dual-basis-riesz}
  \psi_i = \Riesz(\phi^i).
\end{equation}

\subsection{The analysis operator}
\label{ssec:analysis-operator}

\begin{definition}[Analysis operator]
\label{def:analysis-operator}
The \emph{analysis operator} associated with Riesz basis $\{\phi_j\}$ is
\begin{equation}
  \label{eq:analysis-def}
  \Pi : H \to \ell^2, \qquad
  \Pi(m) := \bigl[ (m, \psi_j)_H \bigr]_{j=1}^{\infty}.
\end{equation}
\end{definition}

The analysis operator extracts the coefficient sequence: $\Pi(m) =
\mathbf{m}$. For any Riesz basis it is bounded from $H$ to $\ell^2$.

\subsection{The synthesis operator}
\label{ssec:synthesis-operator}

\begin{definition}[Synthesis operator]
\label{def:synthesis-operator}
The \emph{synthesis operator} is
\begin{equation}
  \label{eq:synthesis-def}
  \pi : \ell^2 \to H, \qquad
  \pi(\mathbf{c}) := \sum_{j=1}^{\infty} c_j\, \phi_j.
\end{equation}
\end{definition}

The synthesis operator reconstructs $m$ from its coefficients. For a
Riesz basis, $\Pi$ and $\pi$ are bounded inverses of each other:
\begin{align}
  \label{eq:pi-Pi-inverses}
  \pi \circ \Pi &= \mathrm{Id}_H, &
  \Pi \circ \pi &= \mathrm{Id}_{\ell^2}.
\end{align}

\begin{codebox}
The coefficient maps correspond to
\[
  \texttt{to\_components}   \equiv \Pi : H \to \ell^2, \qquad
  \texttt{from\_components} \equiv \pi : \ell^2 \to H.
\]
In practice a concrete space has finite \texttt{dim}~$= N$, giving
truncated maps $\Pi_N : H \to \mathbb{R}^N$ and $\pi_N : \mathbb{R}^N \to
H_N$. The operators \texttt{coordinate\_projection} and
\texttt{coordinate\_inclusion} expose $\Pi_N$ and $\pi_N$ as
\texttt{LinearOperator} instances (including their adjoints).
\end{codebox}

\subsection{The dual basis}
\label{ssec:dual-basis}

On the dual space $H'$ the \emph{dual basis} $\{\phi^i\}_{i=1}^\infty$
is defined by $\langle\phi^i, \phi_j\rangle = \delta^i_j$. It forms a
Riesz basis for $H'$. The analysis operator on $H'$ is $\Pi' : H' \to
\ell^2$, $[\Pi'\ell]_j = \langle\ell, \phi_j\rangle$, and the synthesis
operator on $H'$ is $\pi' : \ell^2 \to H'$,
$\pi'(\mathbf{c}) = \sum_j c_j \phi^j$.

The key identity connecting dual and primal bases is:
\begin{equation}
  \label{eq:synthesis-dual-adjoint}
  (\pi')^* = \Pi, \qquad (\Pi')^* = \pi,
\end{equation}
where $(\pi')^* : H \to \ell^2$ and $(\Pi')^* : \ell^2 \to H$ denote the
Hilbert adjoints of $\pi'$ and $\Pi'$ with respect to the inner products
on $H$ and $\ell^2$.

\section{The Gram matrix}
\label{sec:gram-matrix}

\begin{definition}[Gram operator / mass matrix]
\label{def:gram-operator}
Let $\{\phi_j\}$ be a Riesz basis for $H$. The \emph{Gram operator}
(or \emph{mass operator}) is
\begin{equation}
  \label{eq:gram-def}
  \mathbf{M}_\phi := \pi^* \pi : \ell^2 \to \ell^2,
  \qquad
  [\mathbf{M}_\phi]_{jk} = (\phi_k, \phi_j)_H.
\end{equation}
In finite dimensions ($N$-truncation) it becomes the $N \times N$ matrix
\begin{equation}
  \label{eq:gram-finite}
  [\mathbf{M}_{\phi,N}]_{jk} = (\phi_k, \phi_j)_H, \qquad j, k = 1,\ldots,N.
\end{equation}
\end{definition}

\begin{proposition}
\label{prop:gram-spd}
For any Riesz basis, $\mathbf{M}_\phi$ is bounded, self-adjoint, and
strictly positive on $\ell^2$. In particular it is invertible, and its
finite-dimensional truncation $\mathbf{M}_{\phi,N}$ is symmetric positive
definite (SPD) for every $N \geq 1$.
\end{proposition}

\paragraph{Gram matrix and the biorthogonal dual.}
The biorthogonal dual basis is related to the original basis by
\begin{equation}
  \label{eq:biorthogonal-from-gram}
  \psi_j = \sum_i [\mathbf{M}_{\phi}^{-1}]_{ji}\, \phi_i.
\end{equation}
This gives a purely computational recipe: to find $\psi_j$, solve the
linear system $\mathbf{M}_{\phi} \mathbf{c} = \mathbf{e}_j$ (where
$\mathbf{e}_j$ is the $j$-th standard basis vector) and form
$\sum_i c_i \phi_i$.

\paragraph{Analysis via $\phi_j$ alone.}
Substituting \eqref{eq:biorthogonal-from-gram} into the definition of
$\Pi$ gives
\begin{equation}
  \label{eq:analysis-via-phi}
  \Pi = \mathbf{M}_\phi^{-1}\, \pi^* : H \to \ell^2.
\end{equation}
This means the analysis operator can be computed purely from
$\{\phi_j\}$ --- the biorthogonal dual need not be constructed
explicitly. One forms the matrix $\mathbf{M}_{\phi,N}$, inverts it, and
multiplies the vector of inner products $[\langle\phi_j, m\rangle]_j$.

\begin{codebox}
In pygeoinf, \texttt{HilbertSpace.coordinate\_projection} implements
$\Pi_N : H \to \mathbb{R}^N$ via \eqref{eq:analysis-via-phi}:
it evaluates $[\pi_N^*(m)]_j = (\phi_j, m)_H$ and then multiplies by
$\mathbf{M}_{\phi,N}^{-1}$. The Gram matrix itself is available via
\texttt{HilbertSpace.gram\_matrix(N)}.
\end{codebox}

\paragraph{Riesz isomorphism on the coefficient space.}
On the finite coefficient space $\mathbb{R}^N$ (equipped with the norm
inherited from $H$ via $\pi_N$), the Riesz maps are given by the Gram
matrix:
\begin{equation}
  \label{eq:riesz-RN}
  \Riesz_{\mathbb{R}^N}(\mathbf{a}) = \mathbf{M}_{\phi,N}^{-1}\mathbf{a},
  \qquad
  \Riesz_{\mathbb{R}^N}^{-1}(\mathbf{u}) = \mathbf{M}_{\phi,N}\, \mathbf{u}.
\end{equation}
This explains the ubiquity of the Gram matrix in inversion formulas: it
converts coefficient vectors between the primal and dual representations.

\section{Finite-dimensional approximation}
\label{sec:finite-approximation}

All computation in pygeoinf is finite-dimensional. For a Riesz basis
$\{\phi_j\}_{j=1}^\infty$ we fix a truncation parameter $N$ and define:

\begin{equation}
  \label{eq:MN-def}
  H_N := \mathrm{span}\{\phi_1, \ldots, \phi_N\} \subset H.
\end{equation}

The \emph{truncated analysis and synthesis operators} are
\begin{align}
  \label{eq:PiN-piN}
  \Pi_N : H \to \mathbb{R}^N,
  &\quad \Pi_N(m) = \bigl[(m, \psi_j)_H\bigr]_{j=1}^N, \\
  \pi_N : \mathbb{R}^N \to H_N,
  &\quad \pi_N(\mathbf{c}) = \sum_{j=1}^N c_j\, \phi_j.
\end{align}

The \emph{orthogonal projection} onto $H_N$ is
\begin{equation}
  \label{eq:PN-proj}
  P_N := \pi_N \Pi_N : H \to H_N,
\end{equation}
which satisfies $(m - P_N m, \psi_j)_H = 0$ for all $j \leq N$. As $N
\to \infty$, $\|m - P_N m\|_H \to 0$ for every $m \in H$.

\begin{whybox}
The projection $P_N$ is the fundamental link between infinite-dimensional
theory and finite computation. Every algorithm in pygeoinf ultimately
operates on $\mathbb{R}^N$ via the maps $\Pi_N$ and $\pi_N$; the theory
guarantees that as $N$ grows the computed results converge to the
infinite-dimensional truth.
\end{whybox}

\section{Concrete spaces in pygeoinf}
\label{sec:concrete-spaces}

This section introduces the built-in Hilbert space implementations and
explains how to create your own.

\subsection{\texttt{EuclideanSpace}}
\label{ssec:euclidean-space}

$\mathbb{R}^n$ with the standard inner product $(u, v) = \sum_i u_i v_i =
\mathbf{u}^\top \mathbf{v}$. The Riesz map is the identity: $\Riesz = \mathrm{Id}$.
The basis $\{\phi_j\} = \{e_j\}$ is the standard ONB, so $\mathbf{M}_\phi
= \mathbf{I}$ and $\Pi = \pi^* = \mathrm{Id}$.

When you use \texttt{EuclideanSpace(n)}, adjacent operator transposes
\emph{are} adjoints, and you recover the classical formulas of linear
algebra. This is the appropriate choice when variables represent
coefficients in a pre-orthonormalised basis (e.g.\ after a change of
variables $\tilde{m} = \mathbf{L} m$ where $\mathbf{L}$ is a Cholesky
factor of $\mathbf{M}_\phi$).

\subsection{\texttt{MassWeightedHilbertSpace}}
\label{ssec:mass-weighted}

Given a Hilbert space $H$ and a bounded symmetric positive operator
$\mathbf{W} : H \to H'$ (the ``weight operator''), the
\emph{mass-weighted space} $H_{\mathbf{W}}$ is the same underlying vector
space but with inner product
\begin{equation}
  \label{eq:mass-weighted-ip}
  (u, v)_{H_{\mathbf{W}}} = \langle \mathbf{W} u,\, v \rangle.
\end{equation}
The Riesz maps become $\Riesz_{\mathbf{W}} = \Riesz \circ \mathbf{W}$ and
$\Riesz_{\mathbf{W}}^{-1} = \mathbf{W}^{-1} \circ \Riesz^{-1}$.

\paragraph{Motivating use case: discretized $L^2$.}
Using the non-uniform grid inner product from
\cref{sec:why-spaces-matter}, we set $H = \mathbb{R}^N$,
$H_{\mathbf{W}} = \mathbb{R}^N$ with weight $\mathbf{W} = \mathbf{M} =
\mathrm{diag}(\Delta x_j)$. Then $(u, v)_{H_{\mathbf{W}}} = \sum_j
u_j v_j \Delta x_j \approx \int_0^1 u(x) v(x)\,\mathrm{d}x$.

\subsection{\texttt{HilbertSpaceDirectSum}}
\label{ssec:direct-sum-space}

Given Hilbert spaces $H_1$ and $H_2$, their \emph{Hilbert direct sum}
$H_1 \oplus H_2$ consists of pairs $(u_1, u_2)$ with
\begin{equation}
  \label{eq:direct-sum-ip}
  \bigl((u_1, u_2),\, (v_1, v_2)\bigr)_{H_1 \oplus H_2}
  = (u_1, v_1)_{H_1} + (u_2, v_2)_{H_2}.
\end{equation}
The Riesz maps and analysis/synthesis operators decompose blockwise.
This is the natural structure for models that are concatenations of
separately-structured fields (e.g.\ $P$-wave and $S$-wave velocity
perturbations).

\subsection{Implementing your own \texttt{HilbertSpace}}
\label{ssec:custom-space}

To encode a non-standard model space, subclass \texttt{HilbertSpace}.
The abstract interface requires exactly eight methods:

\begin{codebox}
Subclass \texttt{HilbertSpace} and implement:
\begin{itemize}
  \item \texttt{dim} --- integer dimension $N$ of the space.
  \item \texttt{zero()} --- return the zero element.
  \item \texttt{basis\_vector(j)} --- return $\phi_j$ for $j = 0, \ldots, N-1$.
  \item \texttt{inner\_product(x1, x2)} --- compute $(x_1, x_2)_H$.
  \item \texttt{to\_dual(x)} --- compute $\Riesz^{-1}(x) \in H'$.
  \item \texttt{from\_dual(xp)} --- compute $\Riesz(\ell) \in H$ for $\ell \in H'$.
  \item \texttt{to\_components(x)} --- compute $\Pi(x) \in \mathbb{R}^N$.
  \item \texttt{from\_components(c)} --- compute $\pi(\mathbf{c}) \in H$.
\end{itemize}
All other methods (\texttt{norm}, \texttt{distance}, \texttt{project},
\texttt{riesz}, \texttt{gram\_matrix}, \texttt{coordinate\_projection},
\texttt{coordinate\_inclusion}, \ldots) are derived automatically from
the above.
\end{codebox}

\begin{examplebox}
\textbf{Trapezoidal $L^2$ approximation.}
Suppose your model lives in $L^2([0,1])$ and you discretize using a
non-uniform grid $0 = x_0 < x_1 < \cdots < x_N = 1$ with midpoints
$x_j^* = (x_{j-1} + x_j)/2$ and widths $\Delta x_j = x_j - x_{j-1}$.
Implement:
\begin{lstlisting}
class NonUniformL2(HilbertSpace):
    def __init__(self, x):
        self._x = np.asarray(x)        # grid points
        self._dx = np.diff(x)          # cell widths Delta x_j
    @property
    def dim(self):
        return len(self._dx)
    def inner_product(self, u, v):
        return np.dot(u * self._dx, v)  # sum u_j v_j Delta x_j
    def to_dual(self, u):
        return u * self._dx             # (R^-1 u)_j = u_j Delta x_j
    def from_dual(self, up):
        return up / self._dx            # (R ell)_j = ell_j / Delta x_j
    def to_components(self, u):
        return u.copy()                 # basis = piecewise constant
    def from_components(self, c):
        return c.copy()
    def zero(self):
        return np.zeros(self.dim)
    def basis_vector(self, j):
        e = np.zeros(self.dim); e[j] = 1.0; return e
\end{lstlisting}
With this space, \texttt{to\_dual} implements the mass-weighting that
equation~\eqref{eq:correct-min-norm} requires: the adjoint formula
becomes $\mathbf{G}^* = \mathbf{M}^{-1}\mathbf{G}^\top\mathbf{M}_\mathcal{D}$
automatically.
\end{examplebox}

\section{Summary and commutative diagram}
\label{sec:hilbert-summary}

All the maps introduced in this chapter fit together in a single
commutative diagram. Let $H$ be a Hilbert space with Riesz basis
$\{\phi_j\}_{j=1}^N$ (finite-dimensional for conciseness). The spaces,
dual spaces, coefficient space $\mathbb{R}^N$, its dual, and all the maps
between them commute as follows:

\medskip
\begin{center}
\begin{tikzcd}[row sep=3em, column sep=4em]
H
  \arrow[r, blue, "\Pi_N"]
  \arrow[d, "\Riesz^{-1}"']
&
\mathbb{R}^N
  \arrow[d, "\mathbf{M}_{\phi,N}"]
  \arrow[l, bend right=20, blue, "\pi_N"']
\\
H'
  \arrow[r, blue, "\pi_N^*"]
  \arrow[u, "\Riesz"]
&
(\mathbb{R}^N)'
  \arrow[l, bend right=20, blue, "(\Pi_N)^*"']
  \arrow[u, "\mathbf{M}_{\phi,N}^{-1}"]
\end{tikzcd}
\end{center}
\medskip

Reading the diagram:
\begin{itemize}
  \item The \textcolor{blue}{blue horizontal arrows} are the analysis
    ($\Pi_N$) and synthesis ($\pi_N$) operators and their duals.
  \item The vertical arrows on the left are the Riesz maps for $H$.
  \item The vertical arrows on the right are the Gram matrix and its
    inverse, corresponding to the Riesz maps for $\mathbb{R}^N$ equipped
    with the basis inner product.
  \item Every path between two nodes gives the same result: the diagram
    commutes.
\end{itemize}

The outer square expresses $\Riesz^{-1} = \pi_N^* \circ \mathbf{M}_{\phi,N}
\circ \Pi_N$, which is the coordinate representation of the Riesz map.
In \cref{ch:linear-operators} this diagram will be extended to include
a forward operator $G : \modelspace \to \dataspace$ and its adjoint
$G^*$, each appearing as arrows connecting two copies of the above
diagram (one for $\modelspace$, one for $\dataspace$).


% ============================================================================
\chapter{Linear Forms}
\label{ch:linear-forms}
% ============================================================================

\section{Motivation: measurements as linear functionals}
\label{sec:forms-motivation}
% A seismometer, a gravity meter, a magnetometer: each produces a number
% that depends linearly on the model. That's a linear form.

\begin{whybox}
Every single datum in your data vector is the result of applying a
linear functional to the model. Understanding linear forms is
understanding what your data actually \emph{are} mathematically.
\end{whybox}

\section{Definition}
\label{sec:form-definition}
% ℓ: M → ℝ, continuous, linear. Element of M'.

\section{Representation in coordinates}
\label{sec:form-coordinates}
% Components [ℓ(φ₁), ..., ℓ(φ_N)].
% ℓ(m) = ⟨ℓ, m⟩ = cᵀ M⁻¹ m_coeffs.

\section{The Riesz representative (gradient)}
\label{sec:form-riesz-rep}
% gradient = R(ℓ), Hessian = 0. Links to optimisation.

\section{Linear forms as rank-1 operators}
\label{sec:form-as-operator}
% as_linear_operator(): ℓ viewed as M → ℝ.

\begin{codebox}
\texttt{LinearForm} and \texttt{DualVector} represent elements of $\modelspace'$.
Key methods: \texttt{\_\_call\_\_(x)}, \texttt{components}, \texttt{as\_linear\_operator()}.
\end{codebox}


% ============================================================================
\chapter{Linear Operators}
\label{ch:linear-operators}
% ============================================================================

\section{Motivation: the forward operator}
\label{sec:operator-motivation}
% G: M → D maps models to data. This is the mathematical description of
% your forward physics.

\begin{whybox}
The forward operator $G$ is the mathematical encoding of your physics.
Getting its adjoint right is critical for every inversion algorithm ---
and the adjoint depends on the inner products of both spaces, not just
the matrix transpose.
\end{whybox}

\section{Definition and properties}
\label{sec:operator-definition}
% Bounded linear maps. Domain, codomain. Operator algebra: +, scalar ×, @.

\section{The Banach dual (transpose) $G'$}
\label{sec:banach-dual}
% Definition: ⟨G'd', m⟩ = ⟨d', Gm⟩.
% Why this is NOT the adjoint in general.

\section{The Hilbert adjoint $G^*$}
\label{sec:hilbert-adjoint}
% Definition: (Gm, d)_D = (m, G*d)_M.
% The commuting relation: G* = R_M ∘ G' ∘ R_D⁻¹.
% Full derivation.

\begin{pitfallbox}
If you are used to thinking of the matrix transpose as the adjoint, beware:
that is only true in Euclidean space with the standard inner product.
For general Hilbert spaces, $G^* \neq G^\top$. The correct relationship is
$G^* = \Riesz_{\modelspace} \circ G' \circ \Riesz_{\dataspace}^{-1}$.
\end{pitfallbox}

\section{Matrix representations}
\label{sec:matrix-representations}
% Standard matrix: A = Π_D G π_N.
% Galerkin matrix.
% Discrete adjoint: A* = M_φ⁻¹ Aᵀ M_b.

\subsection{The standard (component) matrix}
\label{ssec:standard-matrix}

\subsection{The Galerkin matrix}
\label{ssec:galerkin-matrix}

\subsection{The discrete adjoint formula}
\label{ssec:discrete-adjoint}

\section{Operators as stacks of linear forms}
\label{sec:operator-from-forms}
% G = [ℓ₁, ..., ℓ_M]ᵀ with ℓᵢ ∈ M'.

\begin{codebox}
Given linear forms $\ell_i \in \modelspace'$, construct the forward operator via
\texttt{LinearOperator.from\_linear\_forms([$\ell_1, \ldots, \ell_M$])}.
\end{codebox}

\section{Kernel and range}
\label{sec:kernel-range}
% Null space ker(G), range im(G), fundamental decomposition m = m_null + m_data.

\section{Constructing operators in pygeoinf}
\label{sec:constructing-operators}
% From callables, from matrices, from linear forms.
% Providing vs deriving adjoint/dual mappings.

\begin{codebox}
\texttt{LinearOperator(domain, codomain, mapping)} is the primary constructor.
Optionally provide \texttt{dual\_mapping} ($G'$) or \texttt{adjoint\_mapping} ($G^*$);
the library derives the other via the Riesz maps.
\end{codebox}

\section{The full commutative diagram}
\label{sec:operator-diagram}
% Complete tikz-cd diagram: M, D, M', D', ℝⁿ, ℝ^{N_d}, all maps.


% ============================================================================
\chapter{Nonlinear Operators and Forms}
\label{ch:nonlinear}
% ============================================================================

\section{NonLinearOperator}
\label{sec:nonlinear-operator}
% Interface: __call__, jacobian, adjoint_jacobian.

\section{NonLinearForm}
\label{sec:nonlinear-form}
% Scalar-valued nonlinear functions on Hilbert spaces.

\section{Gradients and Hessians}
\label{sec:frechet-derivatives}
% Fréchet derivatives, connection to optimisation.


% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\part{Geometry and Structure}
\label{part:geometry}
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% ============================================================================
\chapter{Subsets and Convex Sets}
\label{ch:subsets}
% ============================================================================

\section{Why convex sets?}
\label{sec:why-convex}
% Prior information is geometric: the model lies in some set.
% Convexity = computational tractability.

\begin{whybox}
Every piece of prior information you have about your model ---
bounds on parameters, energy constraints, smoothness requirements ---
defines a set of admissible models. Convexity is the mathematical
property that makes these constraints computationally tractable.
\end{whybox}

\section{The Subset interface}
\label{sec:subset-interface}
% is_element(), project(), boundary.

\section{Balls and ellipsoids}
\label{sec:balls-ellipsoids}
% Definition, projection formulas, norm constraints.

\section{Half-spaces and polyhedral sets}
\label{sec:halfspaces-polyhedra}
% Linear inequality constraints. Intersections.

\section{Degenerate sets: EmptySet and FullSpace}
\label{sec:degenerate-sets}

\section{Set operations}
\label{sec:set-operations}
% Intersection, Minkowski sums (via support functions).


% ============================================================================
\chapter{Subspaces}
\label{ch:subspaces}
% ============================================================================

\section{Motivation: equality constraints}
\label{sec:subspace-motivation}
% Linear constraints from conservation laws, boundary conditions.

\section{Orthogonal projectors}
\label{sec:projectors}
% Definition, construction from basis, properties P² = P = P*.

\begin{codebox}
\texttt{OrthogonalProjector.from\_basis(space, [v1, v2, ...])} constructs
the projection $P = \sum_i \hat{v}_i \hat{v}_i^\top$ from spanning vectors.
\end{codebox}

\section{Linear subspaces}
\label{sec:linear-subspaces}

\section{Affine subspaces}
\label{sec:affine-subspaces}
% Translation of linear subspace. Solution set of Bu = w.

\subsection{Construction from tangent bases}
\label{ssec:tangent-basis}

\subsection{Construction from linear equations}
\label{ssec:from-linear-equation}

\section{Tangent spaces and dimension}
\label{sec:tangent-spaces}
% get_tangent_basis(), tolerant Gram--Schmidt.


% ============================================================================
\chapter{Convex Analysis and Support Functions}
\label{ch:convex-analysis}
% ============================================================================

\section{Why convex analysis?}
\label{sec:why-convex-analysis}
% DLI framework: what can we learn without probability?
% Support functions are the key tool.

\begin{whybox}
Convex analysis provides a complete, probability-free framework for
answering the question: given my data and prior constraints, what can I
confidently say about the model? The support function is the central
object that makes this possible.
\end{whybox}

\section{Support functions: intuition}
\label{sec:support-intuition}
% Geometric picture: σ_C(q) = how far C extends in direction q.

\section{Formal definition and properties}
\label{sec:support-definition}
% Convexity, positive homogeneity, subadditivity. Conjugacy.

\section{Subgradients and support points}
\label{sec:subgradients}
% ∂σ_C(q) = argmax, geometric interpretation.

\section{Concrete support functions}
\label{sec:concrete-support}

\subsection{BallSupportFunction}
\label{ssec:ball-support}

\subsection{EllipsoidSupportFunction}
\label{ssec:ellipsoid-support}

\subsection{PolyhedralSupportFunction}
\label{ssec:polyhedral-support}

\subsection{SobolevBallSupportFunction}
\label{ssec:sobolev-support}

\section{Combinators: building complex sets from simple ones}
\label{sec:support-combinators}
% Infimal convolution, sum, scaled sum.

\subsection{Infimal convolution (Minkowski sum)}
\label{ssec:infimal-convolution}

\subsection{Support function sum (intersection)}
\label{ssec:support-sum}

\subsection{Scaled sums}
\label{ssec:scaled-sum}

\section{The dual master equation}
\label{sec:dual-master}
% h_U(q) = inf_λ { ⟨λ, d̃⟩ + σ_B(T'q − G'λ) + σ_V(−λ) }
% Full derivation, geometric interpretation.
% Connection to Al-Attar \& Crawford (2021).


% ============================================================================
\chapter{Gaussian Measures}
\label{ch:gaussian-measures}
% ============================================================================

\section{Why Gaussian measures?}
\label{sec:why-gaussian}
% The natural probabilistic building block for linear-Gaussian problems.

\begin{whybox}
If your prior knowledge about the model is naturally described by a mean
and covariance, and your forward operator is linear, and your measurement
noise is Gaussian, then the full posterior distribution is Gaussian and
can be computed in closed form. This is what \texttt{GaussianMeasure}
provides.
\end{whybox}

\section{Gaussian measures on Hilbert spaces}
\label{sec:gaussian-definition}
% Mean, covariance operator. Cameron--Martin space.
% Difference from finite-dimensional: covariance is an operator.

\section{Sampling}
\label{sec:gaussian-sampling}
% Cholesky-based and randomised approaches.

\section{Push-forward and marginals}
\label{sec:push-forward}
% Am ~ N(Aμ, ACA*).

\section{Conditioning (Bayesian update)}
\label{sec:gaussian-conditioning}
% Posterior formula. Full derivation.


% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\part{Problems and Algorithms}
\label{part:algorithms}
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% ============================================================================
\chapter{Forward Problems}
\label{ch:forward-problems}
% ============================================================================

\section{The forward modelling framework}
\label{sec:forward-framework}
% d = Gm + ε. The forward problem as a data structure.

\section{ForwardProblem and LinearForwardProblem}
\label{sec:forward-classes}
% Constructor, properties, validation.

\begin{codebox}
\texttt{LinearForwardProblem(model\_space, data\_space, forward\_operator,
data\_error\_measure)} bundles the complete forward problem specification.
\end{codebox}


% ============================================================================
\chapter{Deterministic Inversion}
\label{ch:deterministic-inversion}
% ============================================================================

\section{Least-squares inversion}
\label{sec:least-squares}
% min ‖Gm − d‖². Normal equations. Uniqueness.

\section{Minimum-norm inversion}
\label{sec:minimum-norm}
% min ‖m‖ s.t. Gm = d. Pseudo-inverse. Why the norm matters.

\begin{whybox}
The minimum-norm solution depends on the inner product: change the norm,
change the answer. This is why the careful treatment of Hilbert spaces
in \cref{ch:hilbert-spaces} pays off here.
\end{whybox}

\section{Constrained inversions}
\label{sec:constrained-inversions}
% Adding affine constraints Bu = w.

\section{Connection to the dual master equation}
\label{sec:det-inv-dli}
% How deterministic inversions are special cases of DLI.


% ============================================================================
\chapter{Bayesian Inversion}
\label{ch:bayesian-inversion}
% ============================================================================

\section{The Bayesian framework for inverse problems}
\label{sec:bayesian-framework}
% Prior → forward model → likelihood → posterior.
% Stuart (2010) well-posedness.

\section{Linear Bayesian inversion}
\label{sec:linear-bayesian}
% Closed-form posterior. Kalman gain formula.

\begin{codebox}
\texttt{LinearBayesianInversion(forward\_problem, prior)} computes the
posterior Gaussian measure. Access via \texttt{.invert(data)}.
\end{codebox}

\section{Constrained Bayesian inversion}
\label{sec:constrained-bayesian}
% Hard constraints on affine subspaces. Conditional Gaussian.

\section{Posterior uncertainty}
\label{sec:posterior-uncertainty}
% Posterior covariance, marginal variances, confidence ellipsoids.


% ============================================================================
\chapter{Backus--Gilbert and DLI Inference}
\label{ch:backus-gilbert}
% ============================================================================

\section{The inference problem}
\label{sec:inference-problem}
% We want properties p = Tm, not the full model m.

\section{Backus--Gilbert estimators}
\label{sec:bg-estimators}
% Averaging kernels. Resolution--variance trade-off.

\section{The admissible property set $\mathcal{U}$}
\label{sec:admissible-set}
% All property values consistent with data and constraints.
% Characterised by h_U(q).

\section{Computing $\mathcal{U}$ via the dual master equation}
\label{sec:computing-U}
% Numerical evaluation of h_U. Gradient-based optimisation.
% Support points as extreme models.

\section{Polytope outer approximation}
\label{sec:polytope-approximation}
% Three operational regimes: underdetermined, critical, overdetermined.


% ============================================================================
\chapter{Convex Optimisation}
\label{ch:convex-optimisation}
% ============================================================================

\section{Overview of optimisation in pygeoinf}
\label{sec:optim-overview}
% What problems arise, which solvers handle them.

\section{Linear solvers}
\label{sec:linear-solvers}
% CG, MINRES, direct methods.

\section{Convex optimisation algorithms}
\label{sec:convex-algorithms}
% Bundle methods, proximal methods.

\section{Preconditioners}
\label{sec:preconditioners}
% Why preconditioning matters. Available preconditioners.


% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\part{Computational Methods and Visualisation}
\label{part:computation}
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% ============================================================================
\chapter{Computational Utilities}
\label{ch:computational}
% ============================================================================

\section{Randomised matrix decompositions}
\label{sec:randomised}
% Randomised SVD, range finder, Cholesky.

\section{Direct sum operators}
\label{sec:direct-sum-operators}
% Block-diagonal structures.

\section{Parallel computation}
\label{sec:parallel}

\section{Auxiliary functions}
\label{sec:auxiliary}


% ============================================================================
\chapter{Visualisation}
\label{ch:visualisation}
% ============================================================================

\section{The SubspaceSlicePlotter}
\label{sec:slice-plotter}
% Visualising convex sets by slicing.

\section{1D slices}
\label{sec:1d-slices}

\section{2D slices}
\label{sec:2d-slices}

\section{3D slices}
\label{sec:3d-slices}

\section{Worked example: visualising the admissible property set}
\label{sec:vis-example}


% ============================================================================
\appendix
% ============================================================================

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\part*{Appendices}
\addcontentsline{toc}{part}{Appendices}
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% ============================================================================
\chapter{Convex Analysis Primer}
\label{app:convex-primer}
% ============================================================================

\section{Convex sets and functions}
\label{sec:convex-sets-functions}

\section{Separation theorems}
\label{sec:separation}

\section{Conjugate (Fenchel--Legendre) duality}
\label{sec:fenchel-duality}

\section{Subdifferentials}
\label{sec:subdifferentials}

\section{Infimal convolution}
\label{sec:infimal-conv-appendix}


% ============================================================================
\chapter{Functional Analysis Prerequisites}
\label{app:functional-analysis}
% ============================================================================

\section{Banach spaces}
\label{sec:banach-spaces}

\section{Weak and weak-* topologies}
\label{sec:weak-topologies}

\section{Compact operators}
\label{sec:compact-operators}

\section{Spectral theory for self-adjoint operators}
\label{sec:spectral-theory}


% ============================================================================
\chapter{Deferred Proofs}
\label{app:proofs}
% ============================================================================
% Long proofs that would interrupt the main narrative.


% ============================================================================
\chapter{Complete API Quick Reference}
\label{app:api-reference}
% ============================================================================

\section{hilbert\_space module}
\label{sec:api-hilbert}

\section{linear\_forms module}
\label{sec:api-forms}

\section{linear\_operators module}
\label{sec:api-operators}

\section{subsets module}
\label{sec:api-subsets}

\section{subspaces module}
\label{sec:api-subspaces}

\section{convex\_analysis module}
\label{sec:api-convex}

\section{gaussian\_measure module}
\label{sec:api-gaussian}

\section{forward\_problem module}
\label{sec:api-forward}

\section{inversion module}
\label{sec:api-inversion}

\section{linear\_bayesian module}
\label{sec:api-bayesian}

\section{linear\_optimisation module}
\label{sec:api-optimisation}

\section{backus\_gilbert module}
\label{sec:api-bg}

\section{visualisation module}
\label{sec:api-vis}


% ============================================================================
\backmatter
% ============================================================================

% ---------------------------------------------------------------------------
% Bibliography
% ---------------------------------------------------------------------------
\bibliographystyle{plainnat}
\bibliography{bibliography}

% ---------------------------------------------------------------------------
% Index
% ---------------------------------------------------------------------------
% \printindex  % uncomment when index entries are added

\end{document}
